---
title: "Data: Variable Exploration"
author: "Adam Chan, Khanh Chu, Shiyun Chen, Yuanling Zhu, Jing Jing He"
date: "`r format(Sys.time(), '%Y-%m-%d at %H:%M:%S')`"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: 
            collapsed: false
        number_sections: true
        code_folding: show
        theme: united
        highlight: kate
        fig_caption: yes
        encoding: UTF-8
        css: !expr here::here("style.css")
editor_options: 
  chunk_output_type: console
---

```{r section_2_variableExploration, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tidyverse)
library(NbClust)
library(GGally)
library(ggdendro)
library(purrr)
library(stargazer)
library(reactable)
load(here::here("data_retrieval", "process_data.RData"))
```

### Variable Exploration: Descriptive Analysis {.tabset}

#### Numeric Variables

The numeric variables of interest collected are detailed below:

``` {r class.source = 'fold-hide', echo=T, results='hide' }
df_loans %>%
  select(
    "sentiment",
    "lenderRepaymentTerm",
    "loanAmount",
    "daysSinceFundraising",
    "fundedAmount",
    "researchScore"
  ) %>%
  as.data.frame() %>%
  stargazer(type = "html")
```


<div style="text-align: center;">
  <table style="margin: 0 auto; border-collapse: collapse; width: 80%; border-spacing: 10px;">
    <tr>
      <td colspan="6" style="border-bottom: 1px solid black;"></td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">Statistic</td>
      <td style="padding: 10px;">N</td>
      <td style="padding: 10px;">Mean</td>
      <td style="padding: 10px;">St. Dev.</td>
      <td style="padding: 10px;">Min</td>
      <td style="padding: 10px;">Max</td>
    </tr>
    <tr>
      <td colspan="6" style="border-bottom: 1px solid black;"></td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">sentiment</td>
      <td style="padding: 10px;">6,493</td>
      <td style="padding: 10px;">7.280</td>
      <td style="padding: 10px;">4.548</td>
      <td style="padding: 10px;">0</td>
      <td style="padding: 10px;">33</td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">lenderRepaymentTerm</td>
      <td style="padding: 10px;">6,493</td>
      <td style="padding: 10px;">15.096</td>
      <td style="padding: 10px;">6.663</td>
      <td style="padding: 10px;">5</td>
      <td style="padding: 10px;">122</td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">loanAmount</td>
      <td style="padding: 10px;">6,493</td>
      <td style="padding: 10px;">1,427.518</td>
      <td style="padding: 10px;">4,204.432</td>
      <td style="padding: 10px;">100</td>
      <td style="padding: 10px;">300,000</td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">daysSinceFundraising</td>
      <td style="padding: 10px;">6,493</td>
      <td style="padding: 10px;">14.505</td>
      <td style="padding: 10px;">9.077</td>
      <td style="padding: 10px;">0</td>
      <td style="padding: 10px;">42</td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">fundedAmount</td>
      <td style="padding: 10px;">6,493</td>
      <td style="padding: 10px;">0.167</td>
      <td style="padding: 10px;">0.189</td>
      <td style="padding: 10px;">0.000</td>
      <td style="padding: 10px;">1.000</td>
    </tr>
    <tr>
      <td style="text-align: left; padding: 10px;">researchScore</td>
      <td style="padding: 10px;">6,458</td>
      <td style="padding: 10px;">6.926</td>
      <td style="padding: 10px;">4.213</td>
      <td style="padding: 10px;">0.000</td>
      <td style="padding: 10px;">37.500</td>
    </tr>
    <tr>
      <td colspan="6" style="border-bottom: 1px solid black;"></td>
    </tr>
  </table>
</div>

</br>

Note that researchScore here refers to an internal score given by Kiva internally to determine alignment with social impact or effectiveness in achieving positive social outcomes. 

```{r echo=T, warning=FALSE, class.source='fold-hide', results='hide'}
df_loans_selected <- df_loans %>%
  select(
    sentiment,
    lenderRepaymentTerm,
    loanAmount,
    daysSinceFundraising,
    fundedAmount,
    researchScore
  ) %>%
  as.data.frame()

# Reshape the data to long format
df_long <- df_loans_selected %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Create the plot
ggplot(df_long, aes(x = value)) +
  geom_histogram(fill = "#f68060", alpha = .8, bins = 30) +
  labs(title = "Histograms of Selected Variables") +
  theme_minimal() +
  theme_bw() +
  facet_wrap(~ variable, nrow = 2, ncol = 3, scales = "free_x") 
```


#### Country

The top 6 frequent country distribution table is as follows.

```{r class.source = 'fold-hide' , warning=FALSE}
df_loans %>% 
  select(country) %>% 
  count(country) %>% 
  arrange(desc(n)) %>%
  reactable::reactable(
    rownames = FALSE,
    defaultPageSize = 6,
    showPageSizeOptions = F,
    searchable = T,
    compact = T,
    paginateSubRows = T,
    highlight = T,
    bordered = T,
    wrap = T,
    resizable = T
  )

ggplot(df_loans, aes(x = country)) +
    geom_bar(fill = "#f68060", alpha = .8, bins = 30) +
    labs(title = paste("Bar plot of country")) +
    theme_bw() +
    coord_flip()
```


#### Sector

The top 6 frequent sector distribution tables is as follows.

```{r message=FALSE, warning=FALSE, class.source='fold-hide'}
df_loans %>%
  select(sector) %>%
  count(sector) %>%
  arrange(desc(n)) %>%
  reactable::reactable(
    rownames = FALSE,
    defaultPageSize = 6,
    showPageSizeOptions = F,
    searchable = T,
    compact = T,
    paginateSubRows = T,
    highlight = T,
    bordered = T,
    wrap = T,
    resizable = T
  )

ggplot(df_loans, aes(x = sector)) +
  geom_bar(fill = "#f68060",
           alpha = .8,
           width = .4) +
  labs(title = paste("Bar plot of Sector")) +
  theme_minimal() +
  theme_bw() +
  coord_flip() 
```

#### gender

The gender distribution table is as follows. 

```{r message=FALSE, warning=FALSE, class.source='fold-hide'}
df_loans %>%
  select(gender) %>%
  count(gender) %>%
  arrange(desc(n)) %>%
  reactable::reactable(
    rownames = FALSE,
    defaultPageSize = 6,
    showPageSizeOptions = F,
    searchable = T,
    compact = T,
    paginateSubRows = T,
    highlight = T,
    bordered = T,
    wrap = T,
    resizable = T
  )

ggplot(df_loans, aes(x = gender)) +
  labs(title = paste("Bar plot of Gender")) +
  geom_bar(fill = "#f68060",
           alpha = .8,
           width = .4) +
  theme_minimal() +
  theme_bw() +
  coord_flip() 
```

