---
title: "Cluster Analysis: Code"
author: "Adam Chan, Khanh Chu, Shiyun Chen, Yuanling Zhu, Jing Jing He"
date: "`r format(Sys.time(), '%Y-%m-%d at %H:%M:%S')`"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: 
            collapsed: false
        number_sections: true
        code_folding: show
        theme: united
        highlight: kate
        fig_caption: yes
        encoding: UTF-8
        css: !expr here::here("style.css")
editor_options: 
  chunk_output_type: console
---

```{r section_2_clusterAnalysis, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tidyverse)
library(NbClust)
library(GGally)
library(ggdendro)
library(purrr)
library(reactable)
```

The clustering process is depicted in the code chuck below. As seen from the graph, the WSS/TSS ratio flatlines at either 5 or 7 clusters. Given the snapshot of our observation, we will proceed with the smaller number of clusters, which is 5.

```{r, inc-clusterAnalysis, class.source='fold-hide'} 
clusterAnalysis_file_name <- here::here("sections", "section_2", "section_2_clusterAnalysis_clustering.R")
clusterAnalysis_start_line <- 1
clusterAnalysis_end_line <- 72


knitr::read_chunk(
  path = clusterAnalysis_file_name,
  labels = "clusterAnalysis-params",
  from = clusterAnalysis_start_line,
  to = clusterAnalysis_end_line
)

file_in_repo <- stringr::str_sub(
  clusterAnalysis_file_name,
  start = stringr::str_locate(clusterAnalysis_file_name, "/section_2_clusterAnalysis_clustering.R")[1, "start"]
)
``` 

```{r, clusterAnalysis-params, eval = FALSE, echo = TRUE} 
``` 



From here, we will extract summary statistics from these clusters. The numerical means of clusters will be presented in a compact table, while the value count of categorical values is presented cluster-by-cluster. We will add our observations about the clusters along with the categorical count.

```{r}
load(here::here("sections", "section_2", "clusters.Rdata"))

# Function to summarize numerical data and count observations in a cluster
summarize_cluster_numerical <- function(cluster_number, data) {
  numerical_summary <- data %>%
    filter(cluster == cluster_number) %>%
    select(where(is.numeric), -loan_id) %>%
    summarise_all(list(mean = ~ mean(., na.rm = TRUE)))
  
  count_summary <- data %>%
    filter(cluster == cluster_number) %>%
    summarise(count = n()) %>%
    rename(number_of_observations = count)
  
  summary <- bind_cols(numerical_summary, count_summary) %>%
    mutate(cluster = cluster_number)
  
  return(summary)
}

# Generate summaries for clusters 1 to 5
numeric_stats <- map_dfr(1:5, ~ summarize_cluster_numerical(.x, final)) %>%
  pivot_longer(cols = -cluster, names_to = "variable", values_to = "value") %>%
  pivot_wider(names_from = cluster, values_from = value, names_prefix = "cluster_")

print(numeric_stats)
```

General trends within the clusters: Generally they have fairly equal average `fundedAmount` at the time of scraping - regardless of how many days they have been available for donations. This, however, does not indicate an equality between projects: Some are getting funding quicker, while others take more time to reach their current funded amount. A project's research score seems to have a positive relationship with funding speed: The higher that score is, the less time project needs to reach its current funding amount. Sentiment, \# of repayment terms, and loanAmount are on average close between clusters except for one - usually making them the distinguishing factor of that cluster.

```{r}
#Summarize categorical features
summarize_cluster_categorical <- function(cluster_number, data) {
  filtered_data <- data %>%
    filter(cluster == cluster_number) %>%
    select(where(is.factor), -cluster)
  
  summary_data <- lapply(filtered_data, function(col) {
    table_col <- table(col)
    table_col[table_col != 0]  # Remove values that are 0
  })
  
  return(summary_data)
}

cluster_summaries <- list()

for (i in 1:5) {
  cluster_summary <- summarize_cluster_categorical(i, final)
  cluster_summaries[[i]] <- cluster_summary
}

# save(cluster_summaries, file = here::here("sections", "section_2", "ClusterAnalysis_clusterSummaries.RData"))
```

